---
title: "Bayesian Computation in R by Jim Albert"
date: "12/19/2021"
output: 
  html_notebook:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
cat("\014"); rm(list=ls(all=TRUE)); graphics.off()
```


# Introduction to Bayesian thinking
We take the likelihood function as a beta function:
$$
L(p) = p^s(1-p)^f,\ \  0<p<1
$$
and the posterior distribution is proportional to product of prior density and likelihood function:
$$
g(p|data) \propto g(p)L(p)
$$

## Finding posterior distribution using Discrete Prior
We try to find the posterior distribution of the population proportion, p, for the problem given in the book.

```{r}
# Possible prior values
p <- seq(0.05,0.95, by = 0.1)

# Assumed prior distribution for the discrete priors
prior <- c(2,4,8,8,4,2,1,1,1,1)
prior <- prior/sum(prior)

# Quick plot of the prior distribution
plot(p,prior,type = "h",ylab = "Prior Probability")

# Assuming 11 of 27 students sleep at least 8 hours a day
likelihood_func <- function(x){
  return(x^(11)*(1-x)^(27-11))
}


# posterior dist = prior dist * likelihood/sum(prior dist * likelihood)
post <- prior*likelihood_func(p)/sum(prior*likelihood_func(p))

# Combining into a data frame
prior_post <- cbind(p,prior,post)
prior_post

# Quick plot of the posterior distribution
plot(p,post,type = "h",ylab = "Posterior Probability")
```

## Finding posterior distribution using a beta function
We assume that the median and 90^th^ percentiles be 0.3 and 0.5 and the beta density of the prior be a = 3.4 and b = 7.4
```{r}
# Possible prior values
p_beta <- seq(0,1,length = 500)

# density function parameters
a <- 3.4
b <- 7.4

# Assuming 11 of 27 students sleep at least 8 hours a day
s <- 11    # s : success
f <- 16    # f : fail

# Prior distribution
prior_beta <- dbeta(p_beta,a,b)

# Likelihood distribution
like_beta <- dbeta(p_beta,s+1,f+1)

# Posterior distribution
post_beta <- dbeta(p_beta,a+s,b+f)

# Quick plot
plot(p_beta, post_beta, ylab = "Density", lty = 2, lwd = 2, type = "l")
lines(p_beta, like_beta, lty = 1, lwd = 2)
lines(p_beta, prior_beta, lty = 3, lwd = 2)
legend(0.7, 4, c("Prior", "Likelihood", "Posterior"), lty=c(3,1,2), lwd = c(3,3,3) )


```







